Static Analysis by Path Focusing

### 2011-05-02 report/report.tex
minor changes in the report

Static analysis aims at automatically computing properties on programs, such as
the possible values of their variables during execution. This allows to show for
instance that a program will not overflow, will not divide by zero, to compute
loop invariants\dots
The main applications of static analysis are compilation-time optimisation, and
satefy critical systems, such as avionics.

Abstract Interpretation is a general framework used for static analysis.
Linear Relation Analysis (LRA) is a direct application of this framework, that
computes an upper-approximation of the set of the possible states of a numerical
program. The state of a program is defined by the position where we are in the
code, and the actual values of the different numerical variables.
The set of possible states is expressed as a least fixpoint of a set of
equations defining the semantic of the program.

A fundamental fact in static analysis is that it cannot be perfectly precise:
either the set of possible states we compute is unsound (it doesn't contain all
the possible states during execution), either it is incomplete (it contains
states that can never be reached during execution). 
Abstract Interpretation is incomplete, since it always computes an
upper-approximation of the set of possible states.

@@

Satisfiability Modulo Theory (SMT) solving is a field in computer science that
aims to decide the satisfiability of a logic formula containing boolean
predicates and elements of a theory, such as linear arithmetic relations between
integers.
In this report, we present a novel refinment of the Abstract Interpretation
technique, that uses SMT-solving to guide the fixpoint iterations such as it
temporarily focus on a selected path in the control flow graph (the graph
expressing the semantic of the program), in order to
obtain more precise results at reasonable cost.

>>>>

Static analysis aims at automatically computing properties on programs, such as
the possible values of their variables during execution. This allows to show for
instance that a program will not overflow, will not divide by zero, to compute
loop invariants\dots
The main applications of static analysis are compilation-time optimisation, and
the proof of satefy properties in critical systems, such as avionics.

Abstract Interpretation is a general framework used for static analysis.
Linear Relation Analysis (LRA) is a direct application of this framework, that
computes an upper-approximation of the set of the possible states of a numerical
program. The state of a program is defined by the position of the program
counter in the code, and the current values of the different numerical variables.
The set of possible states is expressed as a least fixpoint of a set of
equations defining the semantics of the program.

A fundamental fact in static analysis is that it cannot be perfectly precise:
either the set of possible states we compute is unsound (it doesn't contain all
the possible states during execution), or it is incomplete (it contains
states that can never be reached during execution). 
Abstract Interpretation is sound but incomplete, since it always computes an
upper-approximation of the set of possible states.

@@

Satisfiability Modulo Theory (SMT) solving is a field in computer science that
aims to decide the satisfiability of a logic formula containing Boolean
predicates and elements of a theory, such as linear arithmetic relations between
integers.
In this report, we present a novel refinement of the Abstract Interpretation
technique, that uses SMT-solving to guide the fixpoint iterations, so that it
temporarily focuses on a selected path in the control flow graph (the graph
expressing the semantics of the program), in order to
obtain more precise results at reasonable cost.
